A holomorphic approach to branes in twisted~$M$-theory

### 2022-08-22 fivebranes/fact.tex
merge
update

\section{Factorization algebras in twisted $M$ theory}

@@

It is absolutely crucial for this proposal that we work in a derived setting using the BV formalism. 
The observables of any theory in the BV formalism have the structure of a factorization algebra on spacetime \cite{CG2}. 
Classically, if a theory in the BV formalism is described by a local $L_\infty$ algebra $\cL$ on spacetime $M$ then to an open set $U \subset M$ the factorization algebra of observables assigns the cochain complex
\[
\clie^\bu(\cL(U))
\]
which computes the Lie algebra cohomology of $\cL(U)$. 
The differential on this cochain complex is precisely the Chevellay--Eilenberg differential associated to the~$L_\infty$ structure maps.

@@

We start with the factorization algebra of classical observables of our prototype for the minimal twist of eleven-dimensional supergravity which we recalled in \S \ref{s:twisted}. 
This model is defined on any eleven-manifold $S \times X$ where $X$ is a Calabi--Yau fivefold and $S$ a real oriented one-manifold. 
%In the \S \ref{s:twisted} we recalled a proposal for the description of the minimal twist of eleven-dimensional supergravity on $S \times X$. 
The primary object we used to describe the solutions of the equations of motion was the $\Z/2$ graded local $L_\infty$ algebra $\cL_{sugra}$, see \S \ref{s:Lsugra}. 
%As a graded vector space, observables on an open set $U \subset X \times S$ are thus $\cO(\cL_{sugra}(U)) = \Sym (\Pi \cL_{sugra})^\vee$.
%The differential is encoded by the $L_\infty$ structure which is exactly the Chevalley--Eilenberg differential. 
In terms of this local $L_\infty$ algebra, the $\Z/2$ graded commutative dg algebra of classical observables supported on an open set $U \subset S \times X$ is 
\[
\clie^\bu\left(\cL_{sugra}(U)\right) .
\]
We will denote the classical factorization algebra on $S \times X$ of classical observables by~$\Obs_{sugra}$.

@@

With an eye towards holography for fivebranes we fix a complex three-fold $Z$ where the fivebranes will be supported and consider the eleven-dimensional theory on $\R \times X_Z$ where $X_Z$ is the Calabi--Yau fivefold
\[
X_Z = \text{Tot}(K_Z^{1/2} \otimes \C^2),
\]
the total space of the rank two holomorphic vector bundle $K_Z^{1/2} \otimes \C^2$ over $Z$. 
%In the remainder of this section we denote by $V \to Z$ this rank two holomorphic vector bundle. 
One can also think about $\R \times X_Z$ as the total space of the {\em real} rank five bundle over $Z$; although this description obfuscates the geometric structure used to define the eleven-dimensional theory.
Let 
\[
\pi \colon \R \times X_Z \to Z 
\]
be the map which projects out the real factor followed by the projection to the base of the total space of the complex rank two bundle $p \colon X_Z \to Z$.

Like sheaves, factorization algebras can be pushed forward.
Thus, from the factorization algebra $\Obs_{sugra}$ on $\R \times X_Z$ we obtain a factorization algebra $\pi_* \Obs_{sugra}$ defined on the three-fold $Z$.
Explicitly, to an open set $U \subset Z$ the factorization assigns the cochain complex
\begin{align*}
(\pi_* \Obs_{sugra}) (U) & = \Obs_{sugra} \left(\pi^{-1} U\right) \\
& = \Obs_{sugra}(\R \times p^{-1}U) .
\end{align*}

This factorization algebra is not the factorization algebra associated to an ordinary sort of field theory on $Z$.
Nevertheless there is a subfactorization algebra $\Bar{\pi}_* \Obs_{sugra}$ which admits a natural grading so that each filtered component can be understood as such. 
This grading is determined by looking at an eigenspace decomposition of a certain compact abelian group acting on the fields of the eleven-dimensional theory on $\R \times X_Z$ which we now describe.

One of the primary features of the factorization algebra $\Bar{\pi}_* \Obs_{sugra}$ is that it is of the form $\clie^\bu(\cG)$ where $\cG$ is a sheaf of $L_\infty$ algebras on $Z$ which is presented as the $C^\infty$-sections of an $\infty$-dimensional pro vector bundle.
To get a sense of the construction of this factorization algebra we take a brief excursion to a simpler situation related to branes in five-dimensional Chern--Simons theory.

@@

As we recalled in~\S \ref{??}, Costello's five-dimensional Chern--Simons theory is the result of applying the $\Omega$-background to $M$ theory on a Taub--NUT space.
In this setting, fivebranes which hit the tip of the Taub--NUT are deformed into chiral defects supported along a Riemann surface~$C$ in the five-manifold. 
In analogy with our eleven-dimensional setup, we take the five-manifold to be $\R \times {\rm Tot}(K_C)$ with the deformed fivebranes wrapping the zero section.
Denote by $\Obs_{CS}$ the observables of the semi-classical limit of non-commmutative five-dimensional Chern--Simons theory on $\R \times {\rm Tot}(K_C)$.
Let 
\[\pi \colon \R \times {\rm Tot}(K_C) \to C
\]
be the natural projection and consider the factorization algebra $\pi_* \Obs_{CS}$ on $C$.

@@

For the rank one abelian gauge group, the space of fields of the five-dimensional theory on $\R \times {\rm Tot}(K_C)$ is
\[
\Omega^\bu(\R) \otimes \Omega^{0,\bu}({\rm Tot}(K_C)) [1] .
\]
The interaction is encoded by the symplectic structure on ${\rm Tot}(K_C)$.
%For simplicity, let us assume that the canonical bundle on $C$ is trivializable. 
For $U \subset C$, the complex of observables is given in terms of the Lie algebra cohomology
\[
(\pi_* \Obs_{CS})(U) \simeq \clie^\bu \left(\Omega^{0,\bu}(p^{-1}U)\right)
\]
where $p \colon K_C \to C$.
The Chevalley--Eilenberg complex is taken with respect to the dg Lie algebra structure coming from the holomorphic Poisson structure on the total space of~$K_C$. 
If $U$ is a trivializing cover then
\[
(\pi_* \Obs_{CS})(U) \simeq \clie^\bu \left(\Omega^{0,\bu}(C) \otimes \cO^{hol}(\C)\right) .
\]
On the other hand, we have
\[
(\Bar{\pi}_* \Obs_{CS})(U) \simeq \clie^\bu\left(\Omega^{0,\bu}(C) [[w]] \right) .
\]
Here, $w$ plays the role of a formal coordinate on the fiber of $K_C$ and $w^{k}$ as parametrizes the weight $k$ eigenspace---locally, $w$ is simply the one-form $\d z$.
Using this notation, the embedding \eqref{eqn:taylor1a} is induced by the Taylor series expansion map $\cO(\C) \to \C[[w]]$.
The dg Lie algebra structure on $\Omega^{0,\bu}(C) [[w]]$ is similar to the non-formal case---it is the natural one if we understand $\cO(C)[[w]]$ as functions on the formal neighborhood of the zero section in $K_C$.

@@

Let us return to the situation of a three-fold $Z$ in the eleven-manifold $\R \times \text{Tot}(K_{Z}^{1/2} \otimes \C^2)$. 
The construction of the restricted factorization algebra $\Bar{\pi}_* \Obs_{sugra}$ is completely similar to the simple case of five-dimensional Chern--Simons theory.

@@

Particularly important for us will be the weight zero summand $\cG_Z^{(0)}$.
Again, as a local Lie algebra it is equivalent to one of the form $\Omega^{0,\bu}(Z, \cV^{(0)})$.
When $Z = \C^3$ we will see that the global sections of this $\Z/2$ graded $L_\infty$ algebra is closely related to the exceptional Lie algebra $E(3|6)$ studied by Kac \cite{KacClass}.
%More precisely he weight zero piece is a local Lie algebra enhancement of this exceptional Lie algebra.

@@

\subsection{The membrane decomposition}

We briefly highlight an analogous construction of the restricted factorization algebra describing membranes.
Recall that in the twist we are considering the worldvolume theory of membranes is a three-dimensional partially topological-holomorphic field theory which exist on a product $\R \times C$ where $C$ is a Riemann surface.

We consider the eleven-dimensional theory on $\R \times X_C$ where $X_C$ is the Calabi--Yau fivefold
\[
X_C = \text{Tot}(K_C^{1/4} \otimes \C^4),
\]
the total space of the rank four bundle $K_C^{1/4} \otimes \C^4$ over $C$.
Denote by 
\[
\pi \colon V \to \R \times C
\] 
the pullback of this vector bundle along the projection $\R \times C \to C$.
We can equivalently think of the eleven-manifold as the total space of the bundle $V$ over $\R \times C$. 
In this description, the twisted membranes will be supported along the zero section of $V$. 
The factorization algebra $\Obs_{sugra}$ on ${\rm Tot}(V)$ can be pushed forward along $\pi$ to a factorization algebra $\pi_* \Obs_{sugra}$ on $\R \times C$.

As in the fivebrane case, this is not the factorization algebra associated to an ordinary field theory on $\R \times C$.
Nevertheless there is a subfactorization algebra $\Bar{\pi}_* \Obs_{sugra}$ which admits a natural grading so that each filtered component can be understood as such.

To construct this factorization algebra we consider the action of the compact torus $U(1)^{\times 4}$ which rotates the fibers of the bundle $K_C^{1/4} \otimes \C^4$.
We let $\Bar{\pi}_* \Obs_{sugra}$ be the factorization algebra on $\R \times C$ consisting of observables which are finite sums of integral eigenvectors for this group action.
Again, we find that $\Bar{\pi}_* \Obs_{sugra}$ is of the form $\clie^\bu(\cG_{\R \times C})$ for some $\infty$ sheaf of $L_\infty$ algebras $\cG_{\R \times C}$ on $\R \times C$ which is the sections of a pro vector bundle.

Analogous to the decomposition determined by the $\C^\times$ action considered in \ref{s:cstarfive}, there is a decomposition of $\cG_{\R \times C}$ determined by a (different) $\C^\times$ acting on the fields of the eleven-dimensional theory on ${\rm Tot}(V) \cong \R \times X_C$.

In a local THF chart $U \subset \R \times C$ we can write a generic field of the eleven-dimensional theory on $\pi^{-1} U$ in coordinates as $f(t;z,w)$ where $(t;z) \in U \subset\R \times C$ is THF coordinate and $w=(w_1,w_2,w_3,w_4)$ is the holomorphic fiber coordinate of the rank four bundle $V$.

With this notation in place, we introduce the following $\C^\times$ action on the fields of the eleven-dimensional theory:
\begin{itemize}
\item On the fields $\mu(t;w,z)$ the action is
\[
\lambda \cdot \mu(t;z,w) = \mu(t; z , \lambda w).
\]
\item On the fields $\nu(t;w,z)$ the action is
\[
\lambda \cdot \nu(t;z,w) = \nu(t;z ,\lambda w).
\]
\item On the fields $\beta(t;w,z)$ the action is
\[
\lambda \cdot \beta(t;z,w) = \lambda^{-2} \beta(t; z, \lambda w).
\]
\item On the fields $\gamma(t;w,z)$ the action is
\[
\lambda \cdot \gamma(t;z,w) = \lambda^{-2} \gamma(t; z ,\lambda w).
\]
\end{itemize}

\begin{prop}
The $L_\infty$ structure on $\cL_{sugra}$ is equivariant for this $\C^\times$ action. 
%The odd symplectic form on $\cL_{sugra}$ is weight $-1$ for this $\C^\times$ action. 
\end{prop}

As in the case of fivebranes, this proposition allows us to decompose the local Lie algebra $\cG_{\R \times C}$ into weights. 
Similarly as above one can show that the $n$th eigenspace is quasi-isomorphic to a local Lie algebra of the form
\[
\Omega^\bu(\R) \otimes \Omega^{0,\bu}\left(C, \cV^{(n)}\right)
\]
where $\cV^{(n)}$ is a holomorphic super vector bundle on $C$.

We will state explicit descriptions of the first few eigenspaces without details of the proof. 
The first nontrivial eigenspace is, as in the fivebrane case, $n=-1$. 
Here, one can see that 
\[
\cV^{(-1)} \simeq K_{C}^{1/4} \otimes \C^4 \oplus \Pi K^{3/4}_C \otimes \C^4 .
\]
For $n=0$ one has 
\[
\cV^{(0)} \simeq \T_C \oplus \lie{sl}(4) \otimes \cO_C \oplus \Pi \left(K^{-1/2}_C \otimes \wedge^2 \C^4 \oplus K^{-1/2}_C \otimes S^2 \C^4\right) .
\]
As in the case of fivebranes, the $n=0$ eigenspace is equipped with the structure of a super Lie algebra whose exact form we do not state now.

>>>>

\section{Factorization algebras in twisted $M$ theory}
\label{s:fact}

@@

It is absolutely crucial for this proposal that we work in a derived setting using the Batalin--Vilkovisky (BV) formalism. 
The observables of any theory in the BV formalism have the structure of a factorization algebra on spacetime \cite{CG2}. 
Classically, if a theory in the BV formalism is described by a local $L_\infty$ algebra $\cL$ on spacetime $M$ then to an open set $U \subset M$ the factorization algebra of observables assigns the cochain complex
\[
\clie^\bu(\cL(U))
\]
which computes the Lie algebra cohomology of $\cL(U)$. 
The differential on this cochain complex is precisely the Chevalley--Eilenberg differential associated to the~$L_\infty$ structure maps.

@@

We start with the factorization algebra of classical observables of our prototype for the minimal twist of eleven-dimensional supergravity which we recalled in \S \ref{s:twisted}. 
This model is defined on any eleven-manifold $S \times X$ where $X$ is a Calabi--Yau fivefold and $S$ a real oriented one-manifold. 
%In the \S \ref{s:twisted} we recalled a proposal for the description of the minimal twist of eleven-dimensional supergravity on $S \times X$. 
The primary object we used to describe the solutions of the equations of motion was the $\Z/2$ graded local $L_\infty$ algebra $\cL_{sugra}$, see \S \ref{s:Lsugra}. 
%As a graded vector space, observables on an open set $U \subset X \times S$ are thus $\cO(\cL_{sugra}(U)) = \Sym (\Pi \cL_{sugra})^\vee$.
%The differential is encoded by the $L_\infty$ structure which is exactly the Chevalley--Eilenberg differential. 
In terms of this local $L_\infty$ algebra, the $\Z/2$ graded commutative dg algebra of classical observables supported on an open set $U \subset S \times X$ is 
\beqn\label{eqn:sugraobs}
\clie^\bu\left(\cL_{sugra}(U)\right) .
\eeqn
We will denote the classical factorization algebra on $S \times X$ of classical observables by~$\Obs_{sugra}$.

@@

Fix a complex three-fold $Z$, where the fivebranes will be supported, and assume that it is equipped with a sqaure-root of its canonical bundle $K^{1/2}_Z$.
We consider the eleven-dimensional theory that we introduced in \S \ref{s:twisted} on $\R \times X_Z$ where $X_Z$ is the Calabi--Yau fivefold
\beqn
X_Z = \text{Tot}(K_Z^{1/2} \otimes \C^2).
\eeqn
This is the total space of the rank two holomorphic vector bundle $K_Z^{1/2} \otimes \C^2$ over~$Z$. 
%In the remainder of this section we denote by $V \to Z$ this rank two holomorphic vector bundle. 
One can also think about $\R \times X_Z$ as the total space of the {\em real} rank five bundle over~$Z$; although this description obfuscates the geometric structure used to define the eleven-dimensional theory.
Let 
\beqn
\pi \colon \R \times X_Z \to Z 
\eeqn
be the map which projects out the real factor followed by the projection to the base of the total space of the complex rank two bundle $p \colon X_Z \to Z$.

The classical observables of the eleven-dimensional theory form a factorization algebra on $\R \times X_Z$ that we have denoted by $\Obs_{sugra}$.
Like sheaves, factorization algebras can be pushed forward.
Thus, using $\pi$ we obtain a factorization algebra $\pi_* \Obs_{sugra}$ defined on the three-fold~$Z$.
Explicitly, to an open set $U \subset Z$ the factorization assigns the cochain complex
\begin{align*}
(\pi_* \Obs_{sugra}) (U) & = \Obs_{sugra} \left(\pi^{-1} U\right) \\
& = \Obs_{sugra}(\R \times p^{-1}U) .
\end{align*}

This factorization algebra is not the factorization algebra associated to an ordinary sort of field theory on $Z$.
In physics terminology this feature is due to the presence of an infinite tower of so-called Kaluza--Klein modes which means that the space of fields is not given as the sections of a finite rank vector bundle on~$Z$.
Mathematically, this is due to the fact that the map $\pi$ is not proper.
Nevertheless there is a subfactorization algebra $\Bar{\pi}_* \Obs_{sugra}$ which admits a natural grading so that each filtered component can be understood as such. 
This grading is determined by looking at an eigenspace decomposition of a certain compact abelian group acting on the fields of the eleven-dimensional theory on $\R \times X_Z$ which we now describe.

Like $\Obs_{sugra}$ in equation \eqref{eqn:sugraobs}, the factorization algebra $\Bar{\pi}_* \Obs_{sugra}$ is of the form $\clie^\bu(\cG)$.
Here, $\cG$ is a sheaf of $L_\infty$ algebras on $Z$ which is presented as the $C^\infty$-sections of an $\infty$-dimensional pro vector bundle.
To get a sense of the construction of this factorization algebra we take a brief excursion to a simpler situation related to branes in five-dimensional Chern--Simons theory.

@@

In~\S \ref{s:twistedsugra} we recalled some features of a five-dimensional Chern--Simons type theory that Costello has shown to be the result of applying the $\Omega$-background to $M$-theory on a Taub--NUT space \cite{CostelloM5}.
In this setting fivebranes which hit the tip of the Taub--NUT are deformed into chiral defects supported along a Riemann surface~$C$ in the five-manifold. 
In analogy with our eleven-dimensional setup, we take the five-manifold to be $\R \times {\rm Tot}(K_C)$ with the deformed fivebranes wrapping the zero section.
Denote by $\Obs_{CS}$ the classical observables\footnote{We treat the theory classically as a perturbative theory on $\R \times {\rm Tot}(K_C)$.
We note that the theory is defined in terms of two parameters $\ep, \delta$ where $\delta$ is the Chern--Simons coupling and $\ep$ is the non-commutative parameter controlling the deformation quantization of the holomorphic symplectic manifold ${\rm Tot}(K_C)$.
For convenience we set both $\ep, \delta$ to be one.}
of the non-commmutative five-dimensional Chern--Simons theory on $\R \times {\rm Tot}(K_C)$.
Let 
\beqn
\pi \colon \R \times {\rm Tot}(K_C) \to C
\eeqn
be the natural projection and consider the factorization algebra $\pi_* \Obs_{CS}$ on $C$.

@@

For the rank one abelian gauge group, the space of fields of the five-dimensional theory on $\R \times {\rm Tot}(K_C)$ is
\[
\Omega^\bu(\R) \otimes \Omega^{0,\bu}({\rm Tot}(K_C)) [1] .
\]
The interaction is encoded by the symplectic structure on ${\rm Tot}(K_C)$.
%For simplicity, let us assume that the canonical bundle on $C$ is trivializable. 
For $U \subset C$, the complex of observables is given in terms of the Lie algebra cohomology
\[
(\pi_* \Obs_{CS})(U) \simeq \clie^\bu \left(\Omega^{0,\bu}(p^{-1}U)\right)
\]
where $p \colon K_C \to C$.
The Chevalley--Eilenberg complex is taken with respect to the dg Lie algebra structure coming from the holomorphic Poisson structure on the total space of~$K_C$. 
If $U$ is a trivializing cover then
\[
(\pi_* \Obs_{CS})(U) \simeq \clie^\bu \left(\Omega^{0,\bu}(C) \otimes \cO^{hol}(\C)\right) .
\]
On the other hand, we have
\[
(\Bar{\pi}_* \Obs_{CS})(U) \simeq \clie^\bu\left(\Omega^{0,\bu}(C) [[w]] \right) .
\]
Here, $w$ plays the role of a formal coordinate on the fiber of $K_C$ and $w^{k}$ as parametrizes the weight $k$ eigenspace---locally, $w$ is simply the one-form $\d z$.
Using this notation, the embedding \eqref{eqn:taylor1a} is induced by the Taylor series expansion map $\cO(\C) \to \C[[w]]$.
The dg Lie algebra structure on $\Omega^{0,\bu}(C) [[w]]$ is given by the Moyal commutator. 
This is similar to the non-formal case---it is the natural bracket if we understand $\cO(C)[[w]]$ as functions on the formal neighborhood of the zero section in~$K_C$.

@@

Let us return to the situation of a complex three-fold $Z$ as submanifold the eleven-manifold~$\R \times \text{Tot}(K_{Z}^{1/2} \otimes \C^2)$. 
The construction of the restricted factorization algebra $\Bar{\pi}_* \Obs_{sugra}$ is completely similar to the simple case of five-dimensional Chern--Simons theory as we just discussed.

@@

\brian{make precise statement using $E(3|6)$.}

Particularly important for us will be the weight zero summand $\cG_Z^{(0)}$.
Again, as a local Lie algebra it is equivalent to one of the form $\Omega^{0,\bu}(Z, \cV^{(0)})$.
Since this is the weight zero component, the sheaf of sections is endowed with the structure of a Lie algebra inherited from the local $L_\infty$ algebra $\cL_{sugra}$.
When $Z = \C^3$ we will see that the global sections of this $\Z/2$ graded $L_\infty$ algebra is closely related to the exceptional Lie algebra $E(3|6)$ studied by Kac \cite{KacClass}.
%More precisely he weight zero piece is a local Lie algebra enhancement of this exceptional Lie algebra.

@@

\parsec[s:general_decomp]

\brian{mention $E(3|6)$ module structure}

\begin{prop}
Let $j \geq 1$. 
The complex of vector bundles $\cG^{(j)}$ is quasi-isomorphic to
\beqn
\Omega^{0,\bu}(Z, \cV^{(j)}) 
\eeqn
where $\cV^{(j)}$ is the super holomorphic vector bundle 
\beqn
\begin{tikzcd}
\ul{even} & \ul{odd} \\
S^{j}(\C^2) \otimes \T_Z \otimes K^{-j/2}_Z & S^{j-1}(\C^2) \otimes K^{-(j+1)/2}_Z \\
S^{j+2}(\C^2) \otimes K^{-j/2}_Z & S^{j+1}(\C^2) \otimes \T^*_Z \otimes K^{-(j+1)/2}_Z .
\end{tikzcd}
\eeqn
\end{prop}

\begin{proof}

\begin{itemize}
\item Let $f(w) \in \C[w_1,w_2]_j$ be a homogenous degree $j$ polynomial in the variables $w_1,w_2$ and suppose that $g(z) \del_{z_i}$ is a holomorphic vector field on $Z$. 
Define the $\mu$-type field
\beqn
f(w) g(z) \del_{z_i} + \cdots
\eeqn
\item Let $f(w) \in \C[w_1,w_2]_{j+2}$ be a homogenous degree $j+2$ polynomial and consider the one-form $\del f(w) = \del_{w_1} f(w)\d w_1 + \del_{w_2} f(w) \d w_2$. Using the holomorphic symplectic form we obtain the vector field
\beqn
g(z) \left(\del_{w_1} f(w) \del_{w_2} - \del_{w_2} f(w) \del_{w_1}\right) .
\eeqn
\end{itemize}
\end{proof}
